---
title: "CW7"
editor: visual
---

# Classwork 7

## Question 4

```{r}
holiday_movies <- read.csv("https://bcdanl.github.io/data/holiday_movies.csv")
holiday_movie_genres <- read.csv("https://bcdanl.github.io/data/holiday_movie_genres.csv")
```

## Q4a.
- Provide the R code to generate the data.frame, holiday_movie_with_genres, which combines the two data.frames, holiday_movies and holiday_movie_genres:

```{r}
holiday_movie_with_genres <- holiday_movie_genres %>%
  left_join(holiday_movies, by = c("tconst"))
```

## Q4b. 
- Provide the R code using skimr::skim() to see how the summary statistics—mean, median, standard deviation, minimum, maximum, first and third quartiles—of average_rating and num_votes varies by popular genres and title_type.

  - Consider only the five popular genres, which are selected in terms of the number of titles for each genre.
  - Removes the video type of the titles when calculating the summary statistics.

```{r}
library(dplyr)
library(skimr)
popular_genres <- holiday_movie_genres %>% 
  count(genres) %>% 
  mutate(ranking = dense_rank(-n)) %>% 
  filter(ranking == 5)

unique(holiday_movie_with_genres$title_type)

q4b <- holiday_movie_with_genres %>% 
  filter(genres %in% popular_genres$genres,
         title_type != "video") %>% 
  group_by(genres, title_type) %>% 
  skim(average_rating, num_votes)
```


