--- 
title: "NBA Player Dashboard"
author: Owen Ellick
format: 
  dashboard:
    page-layout: full
    logo: nba.png
    nav-buttons: [github]
    github: https://github.com/OwenEllick/OwenEllick.github.io

---
# Overview

This dashboard provides an interactive overview of NBA player visualizations, and a summary table.

```{r setup}
#| include: false

knitr::opts_chunk$set(
  fig.width = 10,
  fig.height = 6
)

library(tidyverse)
library(ggplot2)
library(gt)

nba_players <- read_csv("http://bcdanl.github.io/data/nba_players.csv")

avg_ppg <- round(mean(nba_players$PTS, na.rm = TRUE),2)
avg_ast <- round(mean(nba_players$AST, na.rm = TRUE),2)
avg_trb <- round(mean(nba_players$TRB, na.rm = TRUE),2)
avg_fgpct <- paste0(round(mean(nba_players$FG_pct, na.rm = TRUE) * 100,2), "%")

nba_avg_points_position <- nba_players %>%
  group_by(Position) %>%
  summarize(Average_Points = mean(PTS, na.rm = TRUE))


```
---

## Value Boxes
```{r}
#| content: valuebox
#| title: "Average Points Per Game"

list(
  icon = "dribbble",
  color = "primary",
  value = avg_ppg
)
```

```{r}
#| content: valuebox
#| title: "Average Assists Per Game"

list(
  icon = "apple",
  color = "success",
  value = avg_ast
)
```

```{r}
#| content: valuebox
#| title: "Average Total Rebounds Per Game"

list(
  icon = "clipboard-data",
  color = "info",
  value = avg_trb
)
```

```{r}
#| content: valuebox
#| title: "Average FG Percentage"

list(
  icon = "percent",
  color = "danger",
  value = avg_fgpct
)

```
## Row {.tabset}
### Visualization: Distribution of Points
```{r}
#| title: Points Distribution
ggplot(nba_players, aes(x = PTS)) +
  geom_histogram(binwidth = 2, fill = "blue", color = "black") +
  labs(
    title = "Distribution of Points per Game",
    x = "Points",
    y = "Frequency"
  ) +
  theme_minimal()

```
### Visualization: Points Per Game vs. Player Efficiency Rating (PER)

```{r}
#| title: Points Per Game vs. PER
ggplot(nba_players, aes(x = PTS, y = PER, color = Position)) +
  geom_point(alpha = 0.7) +
  labs(title = "Points per Game vs. Player Efficiency Rating",
       x = "Points Per Game",
       y = "Player Efficiency Rating (PER)",
       color = "Position") +
  theme_minimal()
```

# Visualizations and Data Table

## {.row}

### Bar Chart: Average Points by Position

```{r}
#| title: Average Points by Position
ggplot(nba_avg_points_position, aes(x = Position, y = Average_Points)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Average Points per Game by Position",
       x = "Position",
       y = "Average Points Per Game") +
  theme_minimal()
```
### Box Plot: Salary by Position

```{r}
#| title: Salary by Position
ggplot(nba_players, aes(x = Position, y = Salary)) +
  geom_boxplot(fill = "tomato", alpha = 0.6, outlier.color = "black") +
  labs(title = "Salary Distribution by Position",
       x = "Position",
       y = "Salary (USD)") +
  theme_minimal()
```
## {.row}

### Scatter Plot: Salary vs. Win Shares

```{r}
#| title: Salary vs. Win Shares
ggplot(nba_players, aes(x = Salary, y = WS, color = Position)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") +
  labs(title = "Salary vs. Win Shares",
       x = "Salary (USD)",
       y = "Win Shares",
       color = "Position") +
  theme_minimal()

```
## {.row}

### Team-Wise Stats Table
```{r}
#| title: Team-Wise Stats Table

nba_players |>
  filter(!str_detect(Team, "/")) %>% 
  group_by(Team) |>
  summarize(
    avg_pts = round(mean(PTS, na.rm = TRUE), 1),
    avg_ast = round(mean(AST, na.rm = TRUE), 1),
    avg_trb = round(mean(TRB, na.rm = TRUE), 1),
    avg_fg_pct = paste0(round(mean(FG_pct, na.rm = TRUE) * 100, 1), "%")
  ) |>
  gt() |>
  tab_header(
    title = "Team-Wise NBA Statistics",
    subtitle = "Average Points, Assists, Rebounds, and FG%"
  ) |>
  cols_label(
    avg_pts = "Avg Points",
    avg_ast = "Avg Assists",
    avg_trb = "Avg Rebounds",
    avg_fg_pct = "Avg FG%"
  ) |>
  tab_options(
    table.font.size = "small",
    table.width = pct(100)
  )
```