---
title: DANL Project
subtitle: "Slam Dunk Salaries: Analyzing NBA Player Earnings"
author: 
  - Owen Ellick
  - Jack Helrigel
  - Jake Starkey
  - Luca Schettino

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

# Introduction

About this project :clap: 
Welcome to our Project, where we dive into the dynamic world of professional basketball through the lens of players salaries and performance statistics. We will analyzing a dataset with NBA Players Salaries and statistics from the 2022-2023 NBA season.

```{r}
library(tidyverse)
library(skimr)
```

# Data

The data.frame `NBAsal` contains a data from the 2022-2023 NBA season. It contains player per-game and advanced statistics for that season with player salary data. This allows us to create a comprehensive resource for understanding the performance and financial aspects of professional basketball players.

## Summary Statistics

```{r}
NBAsal <- read_csv("nba_salaries.csv")
```

```{r}
#| results: asis
#| echo: false
rmarkdown::paged_table(NBAsal) 
```

```{r}
skim(NBAsal) %>% 
  select(-n_missing)
```

## Top Players and Points Per Game

The following shows the top 10 players during the NBA 2022-2023 season based on the number of points per game.

```{r}
top_10_players <- NBAsal %>% 
  group_by(PTS) %>% 
  arrange(desc(PTS)) %>% 
  head(10) %>% 
  relocate(PTS,.before = Position)
  
top_10_players
```

Comments:

-   Joel Embiid, Damian Lillard, Shai Gilgeous-Alexander, Giannis Antetokounmpo, Jayson Tatum, Stephen Curry, Kevin Durant, Lebron James, Donovan Mitchell, and Devin Booker are the top 10 players based on points per game for the 2022-2023 NBA season

## Points Per Game and Salary
Provide both ggplot codes and a couple of sentences to describe the relationship between salary and points per game.

```{r}
ggplot(NBAsal, aes(x= PTS, y= Salary))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "blue")
```
Comments:

* The plot above shows us that during 2022-2023 season players that had a higher average points per game generally had a higher salary than players that had a lower average points per game.

* This implies that players that score more points per game are generally considered more valuable to a franchise which is why they are paid more.

## Top FG%

* Find the ten highest FG% Values of players who have attmpted more than 15 Field Goal Attempts per game

  + FG% is the ratio of field goals made to field goals attempted
  + Generally a higher field goal percentage denotes higher efficiency.
  + A FG% of .500 (50%) or above is considered a good percentage, although it depends on the position.

* Who are the players for those top 10 FG% values?

```{r}
top_FG_Percentage <- NBAsal %>% 
  group_by(`FG%`) %>%
  filter(FGA > 15) %>% 
  arrange(desc(`FG%`)) %>% 
  head(10) %>% 
  relocate(`FG%`, .before = Position) %>% 
  relocate(FGA, .before = Position)

top_FG_Percentage
```

Comments: 

* Zion Williamson, Anthony Davis, Kevin Durant, Giannis Antetokounmpo, Joel Embiid, Kawhi Leonard, De'Aaron Fox, Shai Gilgeous-Alexander, Bradley Beal, Demar DeRozan have the highest FG% for players with more than 15 attempts per game. 

* All these players have a FG% above a .50 which means they are considered highly efficient 

